Panca2Duodenum
===============================================================
Preprocess, prurging of missing data, and explorative analysis
===============================================================

Generated on: `r Sys.time()`

```{r setup}
# Libs
library(Amelia)
library(tidyr)
library(dplyr)
library(ggplot2)
library(nlme)
library(car)
library(psych)

library(ggpubr)
library(rstatix)

#Paths
work_dir <- "/home/bibu/Workspace/panca2duod/"
panca2duoud_raw_fn <- paste(work_dir, "tuep_parsed_final_bela_nov20.csv", sep="")
```

Import raw dataset, check structured column names and create updated wide dataset with structured column names (makers and suffixes for tissue compartment)
```{r}

structured_column_names <- c('pad', '', '', 'halalozas', 'digit', 'ck5_mu',	'ck5_su',	'ck5_bas',	'ck5_lum',	'ck7_mu',	'ck7_su',	'ck7_iv',	'ck7_bas',	'ck7_lum',	'ck20_mu',	'ck20_su',	'ck20_iv',	'ck20_bas',	'ck20_lum',	'ck17_mu',	'ck17_su',	'ck17_iv',	'ck17_bas',	'ck17_lum',	'p63_mu',	'p63_su',	'wt1_mu',	'wt1_su', 'wt1_must',	'wt1_sust',	'ca125_mu',	'ca125_su',	'ca125_iv',	'ca125_bas',	'ca125_lum',	'cdx2_mu',	'cdx2_su',	'cdx2_bas',	'cdx2_lum',	'ca19-9_mu',	'ca19-9_su',	'ca19-9_iv',	'ca19-9_bas',	'ca19-9_lum',	'ecad_mu',	'ecad_su',	'ecad_iv',	'cea-m_mul',	'cea-m_mu',	'cea-m_su',	'cea-m_iv',	'cea-m_bas',	'cea-m_lum',	'muc1_mu',	'muc1_su',	'muc1_iv',	'muc1_bas',	'muc1_lum',	'muc2_mu',	'muc2_su',	'muc2_iv',	'muc2_bas',	'muc2_lum',	'muc5_mul',	'muc5_mu',	'muc5_su',	'muc5_iv',	'muc5_bas',	'muc5_lum',	'muc6_mul',	'muc6_mu',	'muc6_iv',	'muc6_su',	'muc6_bas',	'muc6_lum',	'lym_mu',	'lym_su',	'masp_mu',	'masp_su',	'masp_bas',	'masp_lum',	'masp_iv',	'berep4_mu',	'berep4_su',	'berep4_bas',	'berep4_lum',	'ema_mu',	'ema_su',	'ema_bas',	'ema_lum',	'cd10_mu',	'cd10_su',	'cd10_bas',	'cd10_lum',	'cd10_must',	'cd10_sust',	'pdx1_mu',	'pdx1_su',	'pdx1_bas',	'pdx1_lum',	'p53_mu',	'p53_su',	'smad4_mu',	'smad4_su',	'd2-40_must',	'd2-40_perimu',	'd2-40_sust',	'cd56_mustr',	'cd56_sustr',	'ki67_mu',	'ki67_su',	'ki67_bas',	'ki67_lum',	'polar_mu',	'polar_su',	'field',	'field_pcbil',	'calret_mu',	'calret_su',	'calret_must',	'calret_sust')

panca2duoud_raw <- read.csv(file = panca2duoud_raw_fn, na.strings = "", stringsAsFactors = F, sep= ",")

test_column_names_df  <- rbind(panca2duoud_raw[ 0 , ], structured_column_names)

panca2duoud_wide <- panca2duoud_raw
colnames(panca2duoud_wide) <- structured_column_names

# Filter out columns and rows without useful values. RECHECK indexes in final dataset.
panca2duoud_wide <- panca2duoud_wide[ , -c(2,3,4,5, 122:144)]

```

Key of column suffixes (for tissue compartment)
*mu: mucosal
*su: submucosal
*bas: basal region of the mucosa
*lum:luminal region of the mucosa

iv: intravascular
must: mucosal stroma
sust: submucosal stroma
mul: mucosal cytoplasmic (not relevant)
muf: mucosal surface (relevant similar to mu): fixed for muc5, muc6, ceam
perimu: pericryptal
mustr: mucosal stroma = must
sustr: submucosal stroma = sust
xPolar: lumen formation (0 solid, 1 lumen)
field: field effect (change in normal mucosa, towards pcbil)
pcbil-field: similar to field (not relevant)

Missing data by markers
```{r fig.width=11, fig.height=8}
propmisspercolumn <- function(dataframe) {
  m <- sapply(dataframe, function(x) {
    data.frame(
    nmiss=sum(is.na(x)),
    propmiss=round(mean(is.na(x)),2)
   )
  })
  d <- as.data.frame(t(m))
  d <- sapply(d, unlist)
  d <- as.data.frame(d)
  d$marker <- row.names(d)
  row.names(d) <- NULL
  d <- cbind(d[ncol(d)],d[-ncol(d)])
  return(d[order(d$propmiss), ])
}

dim(panca2duoud_wide)

round(mean(is.na(panca2duoud_wide)),2)

missingpercolumn <- propmisspercolumn(panca2duoud_wide)

print(missingpercolumn)
```

Missing data by rows
```{r}
reportmissperrow <- function(dataframe) {
  pads <- c()
  numMissingValues <- c()
  propMissingValues <- c()
  aRow <- 1
  while(aRow <= nrow(dataframe)) {
    pads <- c(pads, dataframe[aRow, 1])
    numMissingValues <- c(numMissingValues, sum(is.na(dataframe[aRow,])))
    propMissingValues <- c(propMissingValues, round(mean(is.na(dataframe[aRow,])),2))
    aRow <- aRow + 1
  }
  
  missingD <- data.frame(pad = pads, nmissing = numMissingValues, propmissing = propMissingValues)
  
  return(missingD[order(missingD$propmissing), ])
  
  #print(missingD)
  
  #missingD$propmiss_interv <- cut(missingD$propmissing, c(0,10,20,30,40,50,60,70,80,90,100))
  #print(table(missingD$propmiss_interv))
  #return (subset(missingD, propmissing > 50)[, 1])
}

  missingperrow <- reportmissperrow(panca2duoud_wide)
  
  print(missingperrow)
#numCasesMissGreat50 <- reportmissperrow(panca2duoud_wide)
```

Map of missing data
```{r}
missmap(panca2duoud_wide, main = "Map of missing IHC data pcbil2duodenum")
```

Mising values handling strategy, step 1 - remove markers with > 70 % missing values: 
```{r}
markerstotrim <- as.vector(missingpercolumn[missingpercolumn$propmiss > 0.7, ]$marker)
panca2duoud_wide_coltrim <- panca2duoud_wide[, !(names(panca2duoud_wide) %in% markerstotrim)]

dim(panca2duoud_wide_coltrim)
round(mean(is.na(panca2duoud_wide_coltrim)),2)

print(propmisspercolumn(panca2duoud_wide_coltrim))
print(reportmissperrow(panca2duoud_wide_coltrim))

missmap(panca2duoud_wide_coltrim, main = "Map of missing IHC data pcbil2duodenum")
```

Don't remove any case
```{r}
panca2duoud_wide_coltrim_rowtrim <- panca2duoud_wide_coltrim

write.csv(panca2duoud_wide_coltrim_rowtrim, "panca2duod_scores_wide.csv", row.names=FALSE)
```

Reshape filtered dataset from wide to long format
```{r}
gathered_panca2duoud_wide_coltrim_rowtrim <- gather(panca2duoud_wide_coltrim_rowtrim, maker_compart, value, -pad)
tidy_panca2duoud <- separate(gathered_panca2duoud_wide_coltrim_rowtrim, maker_compart, into = c("marker", "compartment"), sep = "_")

tidy_panca2duoud$value <- as.numeric(tidy_panca2duoud$value)

head(tidy_panca2duoud)
tail(tidy_panca2duoud)
unique(tidy_panca2duoud$marker)
unique(tidy_panca2duoud$compartment)
```

```{r, fig.width=20,fig.height=40}
tidy_panca2duoud[is.na(tidy_panca2duoud$compartment), ]$compartment <- "global"
tidy_panca2duoud$compartment <- as.factor(tidy_panca2duoud$compartment)
tidy_panca2duoud$compartment <- factor(tidy_panca2duoud$compartment, levels = c("su", "mu", "bas", "lum", "global", "iv", "mul", "must", "mustr",  "pcbil", "perimu", "sust", "sustr"))

ggplot(subset(tidy_panca2duoud, !is.na(value) & compartment %in% c("mu", "su", "bas", "lum"))) + geom_bar(aes(x=compartment, y=value, fill=compartment),  stat="summary", fun.y=mean) + facet_wrap(~marker, ncol=4)  + theme(strip.text.x = element_text(size = 14, face = "bold")) + theme(axis.text.y = element_text(size = 14))

#ggplot(subset(tidy_panca2duoud, !is.na(value))) + geom_bar(aes(x=compartment, y=value, fill=compartment),  stat="summary", fun.y=mean) + facet_wrap(~marker, ncol=2) + theme(strip.text.x = element_text(size = 14, face = "bold")) + theme(axis.text.y = element_text(size = 14))

write.csv(tidy_panca2duoud, "panca2duod_scores_long.csv", row.names=FALSE)
```



Submucosa - mucosa
```{r, fig.width= 14, fig.asp= 0.85, align = "default"}
# Filter compartments of interest and arrange in a paired df
data_su <- tidy_panca2duoud %>% filter(compartment == "su")
data_mu <- tidy_panca2duoud %>% filter(compartment == "mu")
data_sumu <- rbind(data_su, data_mu)

marker_list_sumu <- data_sumu %>% group_by(marker) %>% summarise(diff_sumu = abs( mean( value[compartment=="su"], na.rm = TRUE) - mean( value[compartment == "mu"], na.rm = TRUE))  ) %>% arrange(desc(diff_sumu)) %>% filter(diff_sumu > 5) %>% pull(marker)
marker_list_sumu <- marker_list_sumu[!marker_list_sumu %in% c("masp", "cd10")]

data_sumu <- data_sumu %>% filter(marker %in% marker_list_sumu)

# Some descriptive stats
sumu_descr <- data_sumu %>%
  group_by(marker, compartment) %>%
  get_summary_stats(value, type = "median")
sumu_descr

# Wilcox test
pwc_sumu <- data_sumu %>%
  group_by(marker) %>%
  wilcox_test(value ~ compartment, paired = TRUE)  %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
pwc_sumu

# Effect sizes
eff_size <- data_sumu  %>%
  group_by(marker) %>%
  wilcox_effsize(value ~ compartment, paired = TRUE)
eff_size

# Box plot
pwc_sumu <- pwc_sumu %>% add_xy_position(x = "compartment")

bxp <- data_sumu %>%
  ggboxplot(x = "compartment", y = "value", color = "compartment", palette = "jco", facet.by = "marker") +
  stat_pvalue_manual(
    pwc_sumu, hide.ns = TRUE, step.group.by = "marker", label = "p = {round(p.adj, 3)}",
    tip.length = 0, step.increase = 0.1)  +
   scale_y_continuous(expand = expansion(mult = c(0, 0.1))) # + ylim (c(0, 110))
bxp
```

```{r, fig.width= 12, fig.asp= 0.75, align = "default"}
# Paired plots
ggpaired(data_sumu, x = "compartment", y = "value", 
         order = c("su", "mu"),
         facet.by = "marker",
         fill = "compartment", palette = "jco", 
        title = "PDAC2Duodenum IHC tissue compartments", 
        ylab = "% positive") + stat_pvalue_manual(pwc_sumu, tip.length = 0, label = "p = {round(p.adj, 3)} {p.adj.signif}", size = 3.7) + expand_limits(y=114) 
```

Mucosa: basal - luminal
```{r, fig.width= 14, fig.asp= 0.85, align = "default"}
# Filter compartments of interest and arrange in a paired df
data_bas <- tidy_panca2duoud %>% filter(compartment == "bas")
data_lum <- tidy_panca2duoud %>% filter(compartment == "lum")
data_baslum <- rbind(data_bas, data_lum)

marker_list_baslum <- data_baslum %>% group_by(marker) %>% summarise(diff_sumu = abs( mean( value[compartment=="bas"], na.rm = TRUE) - mean( value[compartment == "lum"], na.rm = TRUE))  ) %>% arrange(desc(diff_sumu)) %>% filter(diff_sumu > 5) %>% pull(marker)
marker_list_baslum <- marker_list_sumu[!marker_list_sumu %in% c("cd10")]

data_baslum <- data_baslum %>% filter(marker %in% marker_list_baslum)

# Some descriptive stats
baslum_descr <- data_baslum %>%
  group_by(marker, compartment) %>%
  get_summary_stats(value, type = "median")
baslum_descr

# Wilcox test
pwc_baslum <- data_baslum %>%
  group_by(marker) %>%
  wilcox_test(value ~ compartment, paired = TRUE) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
pwc_baslum

# Effect sizes
eff_size_baslum <- data_baslum %>%
  group_by(marker) %>%
  wilcox_effsize(value ~ compartment, paired = TRUE)
eff_size_baslum

# Box plot
pwc_baslum <- pwc_baslum %>% add_xy_position(x = "compartment")

bxp_baslum <- data_baslum %>%
  ggboxplot(x = "compartment", y = "value", color = "compartment", palette = "jco", facet.by = "marker") +
  stat_pvalue_manual(
    pwc_baslum, hide.ns = TRUE, step.group.by = "marker", label = "p = {round(p.adj, 3)}",
    tip.length = 0, step.increase = 0.1
    )  +
   scale_y_continuous(expand = expansion(mult = c(0, 0.1))) # + ylim (c(0, 110))
bxp_baslum
```

```{r, fig.width= 12, fig.asp= 0.75, align = "default"}
# Paired plots
ggpaired(data_baslum, x = "compartment", y = "value", 
         order = c("bas", "lum"),
         facet.by = "marker",
         fill = "compartment", palette = "jco", 
        title = "PDAC2Duodenum IHC tissue compartments", 
        ylab = "% positive") + stat_pvalue_manual(pwc_baslum, tip.length = 0, label = "p = {round(p.adj, 3)} {p.adj.signif}", size = 3.7) + expand_limits(y=114) 
```

Matrix case selection for figures
```{r}
tmp <- panca2duoud_wide_coltrim_rowtrim %>% select("pad", "ck17_su", "ck17_mu", "ck20_su", "ck20_mu", "ca125_su", "ca125_mu", "cdx2_su", "cdx2_mu", "muc2_su", "muc2_mu", "muc5_su", "muc5_mu", "ki67_su", "ki67_mu")
```
